---
title: "p1805_hw1_yc4588"
author: "Yining Cao"
date: "2024-09-19"
output:
  github_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem1

## Load the penguins dataset

```{r}
data("penguins", package = "palmerpenguins")
summary(penguins)
nrow(penguins) # the number of rows
ncol(penguins) # the number of columns
```

* The main variables in the dataset include: 
1. _species_: The species of the penguins, including Adelie(152), Chinstrap(68) and Gentoo(124). 
1. _island_: The islands inhabited by penguins are Biscoe(168), Dream(124) and Torgersen(52). 
1. _bill_length_mm_: The length of the penguin's bill in millimeters(32.10-59.60). 
1. _bill_depth_mm_: The depth of the penguin's bill in millimeters(13.10-21.50). 
1. _flipper_length_mm_: The length of the penguin's flippers in millimeters(172.0-231.0). 
1. _body_mass_g_: The body mass of the penguins in grams(2700-6300).
1. _year_: The year of the penguins, including 2007, 2008 and 2009. 
1. _sex_: The sex of the penguins, including 165 female penguins and 168 male penguins.
* This dataset contains `r nrow(penguins)` rows and `r ncol(penguins)` columns. 
* The mean flipper length for the penguins is `r round(mean(penguins$flipper_length_mm, na.rm = TRUE), digits = 2)` mm.

## Scatterplot of flipper Length(mm) vs bill Length(mm) for penguins

```{r}
library(ggplot2)
p1 = ggplot(na.omit(penguins), aes(x = bill_length_mm, y = flipper_length_mm, color = species)) +
  geom_point() +  
  scale_color_manual(values = c("Adelie" = "#CD9B9B", "Chinstrap" = "#8DB6CD", "Gentoo" = "#8FBC8F")) +
  labs(title = "Flipper Length vs Bill Length",
       x = "Bill Length (mm)",
       y = "Flipper Length (mm)") +
  theme_minimal()
p1

ggsave(filename = "penguins_scatterplot.png", plot = p1, width = 10, height = 7, dpi = 300)
```

* The scatterplot shows the relationship between bill length and flipper length of penguins from different islands.
* Gentoo penguins generally have longer flippers, while Chinstrap and Adelie penguins have shorter bill and flipper lengths.

# Problem2

## Create a data frame

```{r}
library(tibble)
set.seed(123)
df <- tibble(
  number = rnorm(10),
  logical_vector = rnorm(10) > 0,
  character_vector = sample(c("A", "B", "C", "D", "E"), 10, replace = TRUE),
  factor_vector = factor(sample(c("Level1", "Level2", "Level3"), 10, replace = TRUE))
)
```

## Calculate the mean of the number column

```{r}
library(dplyr)
number_mean = df %>% 
  pull(number) %>%
  mean()
number_mean
```
```{r}
logical_vector_mean = df %>% 
  pull(logical_vector) %>%
  mean()
logical_vector_mean
```

```{r}
character_vector_mean = df %>% 
  pull(character_vector) %>%
  mean()
character_vector_mean
```

```{r}
factor_vector_mean = df %>% 
  pull(factor_vector) %>%
  mean()
factor_vector_mean
```
* Means can be calculated for numeric and logical variables.
* Factor and character variables cannot be averaged.

## Convert variables to numeric

```{r  eval = FALSE}
logical_vector = as.numeric(df$logical_vector)
character_vector = as.numeric(df$character_vector)
factor_vector = as.numeric(df$factor_vector)
```
* NA values were introduced due to characters can not be converted to numeric, so characters can not take the mean.
* Factor variable can be converted to numeric without any warning, level1 is converted to 1, level2 is converted to 2, and level3 is converted to 3 so that it can be used in the calculation.

